---
title: "PA1_Template"
author: "Wilco Stel"
output: 
  html_document:
    keep_md: true
---

##part1

### Loading and preprocessing the data

Show any code that is needed to

1. Load the data (i.e. `read.csv()`)

2. Process/transform the data (if necessary) into a format suitable for your analysis



```{r}
library(knitr)
library(dplyr)
library(lubridate)
library(ggplot2)

setwd("C:/Users/Acer E1-731/RepData_Peerassessment1")
data <- read.csv("activity.csv", header = TRUE, sep = ',', colClasses = c("numeric", "character",
                                                                          "integer"))
data$date <- ymd(data$date)
str(data)
head(data)
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print

```

## part2 ### What is mean total number of steps taken per day?

For this part of the assignment, you can ignore the missing values in
the dataset.

1. Make a histogram of the total number of steps taken each day

2. Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
png("instructions_fig/plot1.png")
g<-ggplot(steps, aes(x = steps))
g<- g +  geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
print(g)
dev.off()

mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)

mean_steps

median_steps

```

##part3
In the following we find the daily average activity pattern:

 - Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

 - Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?


```{r, echo=TRUE}

interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))

png('instructions_fig/plot2.png')
g<-ggplot(interval, aes(x=interval, y=steps))
g<-g + geom_line(color = "firebrick")
print(g)
dev.off()



```
##part4
In the following part we will take a look at the imputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Create a new dataset that is equal to the original dataset but with the missing data filled in.

Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r, echo=TRUE}

interval[which.max(interval$steps),]

sum(is.na(data$steps))

data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
sum(is.na(data_full$steps))

steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print

png('instructions_fig/plot3.png')
g<-ggplot(steps_full, aes(x = steps)) 
g<-g +   geom_histogram(fill = "firebrick", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
print(g)
dev.off()

mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)

mean_steps_full

median_steps_full


```
##part5
In the last part we look i there are differences in activity patterns between weekdays and weekends


Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). 



```{r, echo=TRUE}


data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "zaterdag" | weekdays(data_full$date) == "zondag", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)

interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))

png('instructions_fig/plot4.png')
g <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype))
 g<- g + geom_line() 
g<-g +facet_wrap(~weektype, ncol = 1, nrow=2)
print(g)
dev.off()

```
